Application of Machine Learning to Increase the Efficiency of the Global Search Algorithm for Solving Multicriterial Problems

### 2023-03-26 2023/NUMTA/Abstract/Barkalov.tex
full text without references

\title{Title of Your Presentation}
\author{Name1 Surname1, Name2 Surname2}

@@

\institute{University1 of Place1, Address1, Country1 \\
           University2 of Place2, Address2, Country2 \\
\email{Name1@university1.edu, Name2@university2.edu}}

@@

\textbf{Keywords.} Keyword1; keyword2; keyword3.

@@

The goal of the Conference is to create a multidisciplinary round
table for an open discussion on numerical modeling nature by using
traditional and emerging computational paradigms. The Conference
will discuss all aspects of numerical computations and modeling from
foundations and philosophy to advanced numerical techniques. New
technological challenges and fundamental ideas from theoretical
computer science, linguistic, logic, set theory, and philosophy will
meet requirements and new fresh applications from physics,
chemistry, biology, and economy.

Researchers from both theoretical and applied sciences are very
welcome to submit their papers and to use this excellent possibility
to exchange ideas with leading scientists from different research
fields. Papers discussing new computational paradigms, relations
with foundations of mathematics, and their impact on natural
sciences are particularly solicited. A special attention will be
also dedicated to numerical optimization and different issues
related to theory and practice of the usage of infinities and
infinitesimals in numerical computations.

Papers discussing new computational paradigms and their impact on
natural sciences are particularly solicited.

\textit{Please, avoid figures and tables in the text. Max 4
bibliographic references are allowed.}

@@

This research was supported by the following grants...

@@

\bibitem{Proceedings} Author I., Author I.I. (2019) An
abstract. In Proceedings of the \emph{International
Conference``XXX''} (ed. by Editor E.), Place (Country), p.~1.

\bibitem{BookChapter} Author X. (2019) A
chapter. In \emph{YYY Book} (ed. by Editor E.), Place (Country),
pp.~10--20.

>>>>

\title{Application of Machine Learning to Increase the Efficiency of the Global Search Algorithm for Solving Multicriterial Problems}
\author{Konstantin Barkalov, Vladimir Grishagin, Evgeny Kozinov}

@@

\institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \\
\email{\{konstantin.barkalov,evgeny.kozinov\}@itmm.unn.ru}, vagris@unn.ru}

@@

\textbf{Keywords.} Multicriterial problems; global optimization; logistic regression. % machine learning;

@@

Decision making models described as problems of multicriterial optimization are very complicated for investigation because of the property of criteria contradictoriness that leads to the notion of the solution as the set non-dominated parameters (Pareto set). The complexity of these problems increases significantly in the case where criteria are multiextremal. In the approaches based on convolution schemes reducing the multicriterial problem to a family of scalar (single criterion) optimization subproblems, these subproblems are multiextremal as well. This circumstance requires the use of global optimization methods to solve those.

In the paper a novel algorithm for multicriterial black-box optimization with multiextremal criteria are considered. This algorithm applies ideas of complexity reduction when the initial multicriterial problem are reduced to a set of scalar subproblems on the base of maximum convolution, then each scalar multidimensional subproblem is transformed to equivalent one-dimensional problem by means of Peano mapping which provides single-valued correspondence between N-dimensional hyperparallelepiped and the unit interval of the real axis. For solving univariate problems a global optimization algorithm with guaranteed convergence to global optimum is used. As a core novel feature, this algorithm includes in its computational scheme a machine learning procedure with combination of accumulating the information of solved subproblems that allows one to significantly accelerate the building the Pareto set.

For verification of workability and estimation of efficiency, the representative computational experiment on test sets of multicriterial problems with multiextremal criteria constructed on the base test class GKLS  has been conducted for different dimensionalities and number of criteria.

Along with the proposed algorithm, in the experiment some known method belonging to the class of genetic algorithms have been tested and compared where the hypervolume index  HV was used for assessment of algorithms' quality. According to the results of the experiment, the proposed algorithm  has demonstrated its essential advantage over other methods.

@@

This research was supported by the Russian Science Foundation, project No. 21-11-00204.

@@
